\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2018-07-31 sections/methods.tex
Fix indentation of items in methods
Change section heading to "Materials and Methods"
Add datasets subsection and reformat methods

\section*{Methods}

  The flowchart describing the workflow of the \ac{mind} pipeline is as shown in Figure \ref{fig:mind_pipeline}.
  The pipeline seamlessly integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
  The input to the pipeline by default is the raw 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
  The final output of the pipeline is the inferred network of co-occurrence associations among the microbes present in the samples.
  In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
  We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
  We propose to use mock synthetic data simulated using an Illumina read-simulator \cite{Escalona2016} and a mock community data-set \cite{Bokulich2016} as standards for comparing the counts matrices and the assigned taxonomies.
  We will then use a voting method to obtain a co-occurrence network that has links predicted by multiple network inference methods.

@@

  \subsubsection*{Denoising and Clustering}
    \vspace{-5mm}
    This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
    It incorporates three processes: quality control, denoise/cluster and chimera checking.
    \begin{itemize}
    \item The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
    \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables \footnote{Some methods perform clustering and taxonomy assignment in the same step}.
    \item The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
    \end{itemize}
    The output of this module is a counts matrix that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix). The pipeline incorporates \ac{qiime1} \cite{Caporaso2010}, \ac{qiime2}, \ac{dada2} \cite{Callahan2016} and Deblur \cite{Amir2017} tools in this module.

@@

  \subsubsection*{OTU and ESV Processing}
    \vspace{-5mm}
    This module deals with normalization, filtering and applying transformations on the \ac{otu} or \ac{esv} counts matrix.
    \begin{itemize}
    \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth \footnote{The total number of reads obtained from a particular sample} in different samples. This is commonly performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
    \item Filtering is usually performed to remove samples or features from the count matrix that are sparse.
    \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix.
    \end{itemize}

@@

  \subsubsection*{MIND Data Visualizer}

>>>>

\section*{Materials and Methods}

  \subsection*{Datasets}

    \vspace{-5mm}

    The mock community 16S datasets was obtained from \cite{Bokulich2016}.

    The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} obtained from \cite{Kang2017}.

    Simulated data was generated using Grinder \cite{Angly2012}.

  \subsection*{16S processing pipeline}

    \vspace{-5mm}

    The flowchart describing the workflow of the \ac{mind} pipeline is as shown in Figure \ref{fig:mind_pipeline}.
    The pipeline integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
    The input to the pipeline by default is the raw community 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
    The final output of the pipeline is the inferred network of co-occurrence associations among the microbes present in the samples.
    In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
    We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
    We propose to use mock synthetic data simulated using an Illumina read-simulator \cite{Escalona2016} and a mock community data-set \cite{Bokulich2016} as standards for comparing the counts matrices and the assigned taxonomies.
    We will then use a voting method to obtain a co-occurrence network that has links predicted by multiple network inference methods.

@@

    % TODO: Talk about what is happening in the pipeline NOT what that steps does in theory
    \subsubsection*{Denoising and Clustering}
      \vspace{-5mm}
      This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
      It incorporates three processes: quality control, denoise/cluster and chimera checking.
      \begin{itemize}
        \item The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
        \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables \footnote{Some methods perform clustering and taxonomy assignment in the same step}.
        \item The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
      \end{itemize}
      The output of this module is a counts matrix that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix). The pipeline incorporates \ac{qiime1} \cite{Caporaso2010}, \ac{qiime2}, \ac{dada2} \cite{Callahan2016} and Deblur \cite{Amir2017} tools in this module.

@@

    \subsubsection*{OTU and ESV Processing}
      \vspace{-5mm}
      This module deals with normalization, filtering and applying transformations on the \ac{otu} or \ac{esv} counts matrix.
      \begin{itemize}
        \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples. This is commonly performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
        \item Filtering is usually performed to remove samples or features from the count matrix that are sparse.
        \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix.
      \end{itemize}

@@

  \subsection*{Statistical Analysis}

    Tests

  \subsection*{MIND Data Visualizer}
